Newly Left-Behind Children's Educational Performance Drops Slightly After Parental Migration

### 2023-03-23 MScThesisText.tex
Adjust sampling and reference entries

The internal migration stemmed from the neoliberal economic regime. Scholars like Duckett (2020) and Zhang (2018) argued that China conforms to the neoliberal trend, extending the logic of the market to other domains of human life. In the 1980s, the ruling communist party ditched the planned economy, initiated market reforms, and opened up the country. Since then, China has integrated in the global economy and contributed substantial growth to it. The economic reform and globalization have enriched the coastal cities like Shanghai and Shenzhen. The rural hinterlands, however, have lagged behind. Simultaneously, many people have left the countryside for the city. The regional inequality needs social policy to amend. But China's welfare program, as Qian (2021) argued, is fragmented and stratified. People who reap the greatest benefits are public servants, urban dwellers, and formally employed workers. As rural migrants arrive in the city, they find themselves with limited access to social welfare, such as health care and children's education (Qian, 2021). The welfare system's exclusion of rural migrants means that the state shifts its responsibility to individuals, resembling the neoliberal governmentality (Zhan, 2020).

@@

In other words, the hukou system binds one's citizenship rights with one's place of registry. In theory, Chinese citizens can migrate freely within the country. In practice, they cannot change hukou freely. When rural residents move to the city, it takes months -- if not years -- to apply for a new hukou. Before receiving urban hukou, they can hardly send their children into public schools in the host city (Wan \& Vickers, 2022). Migrant workers have organized petitions, strikes, and protests. These actions often fell on deaf ears and sometimes even ended with police crackdown (Chan, 2019). Facing financial and policy constraints, many migrant laborers have to leave children in their hometowns. Only during the winter and summer breaks can these children reunite with parents.

@@

Social scientists have researched extensively about the education of Chinese left-behind children, but the current literature still has three major shortcomings. The first shortcoming lies in sampling. For many studies, the geographic coverage is limited to one to a few counties; The participants are children left behind in rural areas but not those in urban areas (see Hu, 2019; Zhou et al., 2014). S. Wang (2023) noted that the number of urban left-behind children have been growing, and the trend is expected to continue. The second shortcoming is the usage of cross-sectional data (see Huang \& Gong, 2022; Y. Liu et al., 2022). Wassink and Viera (2021) argued that cross-sectional surveys merely focus on single time points, thereby failing to capture the dynamic nature of migratory journey. Research results would be under- or over-estimated ``depending on the timing of data collection'' (p.~1766). The third shortcoming lies in the measurement of outcome. Chang et al. (2019) and L. Wang et al. (2019) measured students' educational performance with the standardized test score of only one academic subject, namely mathematics. To tackle these limitations, I adopted a nationally representative two-wave panel dataset. My sample includes urban left-behind children as well. I included the test results of multiple subjects to measure comprehensive academic achievement.

@@

Between parental migration and children's education, the mechanisms trace to two theoretical frameworks. One is the resource generation model. It holds that parental migration brings forth greater resources in the form of economic capital (Brauw, 2019). Economic capital is, in Bourdieu's words, ``immediately and directly convertible into money and may be institutionalized in the form of property rights'' (2002, p. 16). Economic resources are transferable to human capital, which is the stock of ``skills, talents, health, and expertise'' in people that enhance their productivity (Botev et al., 2019; Goldin, 2016). Yang (2008) noted that in the Philippines, a sharp depreciation of peso increased the amount of remittances from migrants working abroad. This subsequently boosted children's school enrollment and educational expenditure. Cebotari (2020) developed a fixed effect model with longitudinal data collected in Ghana. He found that stay-behind children's education benefited from ``the steady inflow of both in-kind and monetary remittances'' contributed by migrant parents. In China, Gu (2022) interviewed migrant workers and their family members. She found that they aspire to move up the social ladder; Such aspiration translates into ``intensive and all-out investment in children's education'' (p.~526). Oftentimes, labor migration presented the only way to sustain their children's education and livelihood. Peng (2018) interviewed first-generation migrant mothers whose initial migration dated back to the 1990s. To them, ``good mothering'' primarily meant offering economic support for their children in education, which can help their children climb the ladder of success.

The other theoretical model is the family disruption model. Children flourish under parental nurturing care. This term refers to the ``stable environments that promote health and adequate nutrition, protect from threats \ldots{} opportunities for learning,'' and ``responsive, emotionally supportive and developmentally enriching relationships'' (Black et al., 2021, p. 1). Parental migration decreases nurturing care, as characterized by less homework supervision and emotional support by parents. Hong talked to a group of left-behind children at a rural school (Hong \& Fuller, 2019). These students expressed the utter sense of loneliness ``with no advocate, no support and with no one looking out for them'' (p.~13). Many of them felt indifferent to grades and disengaged with school. Some even longed to quit school to pursue a free, independent life (p.~14). Similarly, Murphy (2022) talked to left-behind adolescents at a village. Some of them ``embraced the `left-behind' label'' because it captured their need for more parental care (p.~185), ``especially companionship and affection in daily life'' (p.~186). According to C. Chen (2023), caregiver-child interactions are infrequent and unstable for left-behind children. She showed that the loss in such interactions ``mediates almost all of the adverse effects of parental absence.'' Her conclusion was based on structural equation modeling using a cross-sectional survey with a sample of Chinese children aged three to fifteen.

@@

I obtained the data from the China Education Panel Survey (CEPS) (NSRC, n.d.). It is a nationally representative, school-based survey featuring junior high school students. Currently, two waves of data are available. The CEPS team operates under the National Survey Research Center at the Renmin University of China. The survey team adopted multi-stage stratified probability proportional to size sampling and administered a paper-and-pencil questionnaire to each participant. The wave one baseline survey took place in the 2013-2014 academic year with a sample size of about twenty thousand students in the seventh and ninth grades. They were nested in 438 classrooms of 112 schools in 28 urban districts or rural counties (hereafter `counties'). The wave two follow-up survey in the 2014-2015 academic year lost track of the ninth grade cohort and only surveyed 10279 of the seventh grade cohort. I further restricted the sample to students living with both parents at the original domicile (place of hukou registry) at the baseline survey. To tackle missing values, I adopted list-wise deletion for the dependent, independent, and control variables. The final sample size amounts to 4062.

@@

The control variables were all measured in the baseline survey. Covariates at children's level included age, gender, self-rated health, hukou type, number of siblings, whether the child is ethnic minority or Han Chinese, whether the child had skipped grade or repeated grade, and the level of education that the child aspires to achieve. At the household level, I controlled the levels of education of father and mother, the parental expectation on the child's level of education, whether the parent has a white collar job, whether the household has internet access, the perceived household financial conditions, and the number of extracurricular books at home. The lists of covariates are in Table~\ref{tab:tab-ctvar-dit} -- Table~\ref{tab:tab-ctvar-msr} in the Results section.

@@

I built a two-way fixed effect regression model in a difference-in-differences (DID) design, utilizing parental migration as the treatment (Bai et al., 2018). The treatment group consists of students left behind by migrating parents at a certain point of time between the two waves of survey, and the control group includes those who do not. I compared the treatment group's pre- and post-treatment outcomes in relation to the control group.

Below is my first version of DID model: \[\Delta score_{i,j} = \alpha + \beta \cdot migr_{i} + \gamma \cdot score_{i;1} + \lambda \cdot S_{j} + \varepsilon_{i,j}\] It is unrestricted because it does not restrict on the coefficient associated with the baseline scores. The model is unadjusted as it does not adjust for additional covariates. Theoretically, it is unnecessary to include covariates that vary over group but remain constant over time; They would cancel out in the two-way fixed effect model (Huntington-Klein, 2022). For student \(i\) in school \(j\), \(\Delta score_{i,j}\) is the change in test score between the two waves of surveys, \(migr_{i}\) is the treatment dummy variable, \(score_{i,j;1}\) is the test score at baseline, and \(S_{j}\) is the school fixed-effect. School fixed-effect can account for differences due to time-invariant factors at the school and county levels, as students nested in one school are also in the same county. The standard errors are clustered at the school level. An alternative is to use a class fixed-effect to control for time-invariant factors at the classroom level and above. However, about one-sixth of the students were reassigned into new classes at the end of baseline academic year. As the reassignment of classroom will complicate the model, I ruled out the classroom fixed-effect.

@@

I adopted \texttt{R} (version 4.2.2) for all the analyses (R Core Team, 2022). This involved the following packages: \texttt{tidyverse} (version 2.0.0) and \texttt{haven} (version 2.5.0) for data cleaning (Wickham et al., 2019, 2022); \texttt{fixest} (version 0.11.1) for econometric modeling (Berge et al., 2022; Berge, 2018); \texttt{MatchIt} (version 4.5.1) and \texttt{cobalt} (version 4.3.2) for matching (Greifer, 2022; Ho et al., 2022, 2011); \texttt{modelsummary} (version 1.3.0), \texttt{vtable} (version 1.4.1), and \texttt{kableExtra} (version 1.3.4) for presenting the results (Arel-Bundock, 2022; Huntington-Klein, 2021; Zhu, 2021); and \texttt{papaja} (version 0.1.1) for typesetting (Aust \& Barth, 2022).

@@

Table~\ref{tab:tab-idpvar} illustrates the independent variables, namely the migration status reported in the wave two survey. All 4062 children were living with parents in one household in baseline, but about sixteen percent saw at least one parent migrated between the two waves of survey. I call these students ``newly left-behind children'' and their households ``new migration households.'' In these new migration households, only-mother migration and only-father migration each accounted for two out of five cases. The remaining one-fifth were both-parents migration.

@@

\begin{longtable}[]{@{}lll@{}}
\caption{\label{tab:tab-idpvar}Independent variables by parental migration}\tabularnewline
\toprule()
Variable & N & Percent \\
\midrule()
\endfirsthead
\toprule()
Variable & N & Percent \\
\midrule()
\endhead
Parental Migration & & \\
\ldots{} No & 3422 & 84.2\% \\
\ldots{} Yes & 640 & 15.8\% \\
Parental Migration Type & & \\
\ldots{} Both Parents at Home & 3422 & 84.2\% \\
\ldots{} Both Parents Migrated & 130 & 3.2\% \\
\ldots{} Only Father Migrated & 252 & 6.2\% \\
\ldots{} Only Mother Migrated & 258 & 6.4\% \\
\bottomrule()
\end{longtable}

@@

\caption{\label{tab:tab-dpvar}Dependent variables by parental migration}
\centering
\begin{tabular}[t]{llllll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Parental Migration} \\
\cmidrule(l{3pt}r{3pt}){2-5}
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{No = 3422} & \multicolumn{2}{c}{Yes = 640} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
Variable & Mean & Sd & Mean & Sd & Test\\
\midrule
Cognitive test score difference & 11.48 & 19.85 & 9.58 & 20.71 & p=0.028*\\
\hspace{1em}Wave 1 score & 56.75 & 16.63 & 55.98 & 16.62 & p=0.282\\
\hspace{1em}Wave 2 score & 68.22 & 18.45 & 65.56 & 19.54 & p<0.001***\\
Academic exam score difference & -2.78 & 10.91 & -4 & 11.78 & p=0.01**\\
\hspace{1em}Wave 1 score & 71.21 & 16.37 & 70.4 & 16.93 & p=0.256\\
\addlinespace
\hspace{1em}Wave 2 score & 68.43 & 18.28 & 66.4 & 19.49 & p=0.011*\\
\bottomrule
\multicolumn{6}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{6}{l}{\rule{0pt}{1em}* p < 0.05, ** p < 0.01, *** p < 0.001}\\
\end{tabular}
\end{table}

@@

\caption{\label{tab:tab-ctvar-dit}Dichotomous control variables by parental migration}
\centering
\begin{tabular}[t]{llllll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Parental Migration} \\
\cmidrule(l{3pt}r{3pt}){2-5}
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{No = 3422} & \multicolumn{2}{c}{Yes = 640} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
Variable & N & Percent & N & Percent & Test\\
\midrule
child.is.girl &  &  &  &  & X2=12.261***\\
... 0 & 1637 & 48\% & 355 & 55\% & \\
... 1 & 1785 & 52\% & 285 & 45\% & \\
child.has.rural.hukou &  &  &  &  & X2=1.09\\
... 0 & 1887 & 55\% & 338 & 53\% & \\
\addlinespace
... 1 & 1535 & 45\% & 302 & 47\% & \\
child.in.boarding.school &  &  &  &  & X2=4.49*\\
... 0 & 2470 & 72\% & 435 & 68\% & \\
... 1 & 952 & 28\% & 205 & 32\% & \\
child.is.ethnic.minority &  &  &  &  & X2=0.514\\
\addlinespace
... 0 & 3190 & 93\% & 591 & 92\% & \\
... 1 & 232 & 7\% & 49 & 8\% & \\
child.had.skipped.grade &  &  &  &  & X2=0.056\\
... 0 & 3386 & 99\% & 632 & 99\% & \\
... 1 & 36 & 1\% & 8 & 1\% & \\
\addlinespace
child.had.repeated.grade &  &  &  &  & X2=10.485**\\
... 0 & 3121 & 91\% & 557 & 87\% & \\
... 1 & 301 & 9\% & 83 & 13\% & \\
child.went.to.preschool &  &  &  &  & X2=0.923\\
... 0 & 523 & 15\% & 108 & 17\% & \\
\addlinespace
... 1 & 2899 & 85\% & 532 & 83\% & \\
parent.has.white.collar.job &  &  &  &  & X2=0.001\\
... 0 & 2705 & 79\% & 505 & 79\% & \\
... 1 & 717 & 21\% & 135 & 21\% & \\
home.has.internet &  &  &  &  & X2=12.967***\\
\addlinespace
... 0 & 1054 & 31\% & 244 & 38\% & \\
... 1 & 2368 & 69\% & 396 & 62\% & \\
\bottomrule
\multicolumn{6}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{6}{l}{\rule{0pt}{1em}0 = No, 1 = Yes; * p < 0.05, ** p < 0.01, *** p < 0.001}\\
\end{tabular}
\end{table}

@@

\caption{\label{tab:tab-ctvar-cnt}Categorical and continuous control variables by parental migration}
\centering
\begin{tabular}[t]{llllll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Parental Migration} \\
\cmidrule(l{3pt}r{3pt}){2-5}
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{No = 3422} & \multicolumn{2}{c}{Yes = 640} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
Variable & Mean & Sd & Mean & Sd & Test\\
\midrule
child.age & 12.45 & 0.63 & 12.51 & 0.73 & F=4.263*\\
child.number.of.sibling & 0.58 & 0.74 & 0.66 & 0.82 & F=4.896*\\
child.health & 4.17 & 0.87 & 4.14 & 0.9 & F=0.576\\
child.education.aspiration & 3.96 & 1.12 & 3.93 & 1.19 & F=0.522\\
mother.education.level & 1.53 & 1.17 & 1.57 & 1.25 & F=0.468\\
\addlinespace
father.education.level & 1.72 & 1.18 & 1.75 & 1.21 & F=0.274\\
parent.expectation & 4.19 & 0.91 & 4.14 & 0.95 & F=1.627\\
home.extracurricular.book & 3.42 & 1.18 & 3.32 & 1.28 & F=3.392\\
home.economic.resource & 2.87 & 0.55 & 2.84 & 0.57 & F=1.52\\
\bottomrule
\multicolumn{6}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{6}{l}{\rule{0pt}{1em}* p < 0.05, ** p < 0.01, *** p < 0.001}\\
\end{tabular}
\end{table}

@@

I first explain the results on children's cognitive test score. The unadjusted DID models show that newly left-behind adolescents are more likely to score lower in cognitive tests than do their non-left-behind counterparts (see Table 6 below). This lends support to Hypothesis 1a. Parental migration is linked to a 2.56 points decrease in cognitive test score. Hypothesis 2a is only partially supported. Compared with the non-left-behind peers, children with only father absent are associated with greater decrease (2.87 points) in cognitive test score than those with only mother absent (2.15 points). The adjusted DID models produce results similar to the unadjusted models, with one exception (see Table A1 in the appendix). That is, for the cognitive score, the both-parent-migrated coefficient's magnitude is lower than that of father-migrated.

@@

Parental migration is not a randomly assigned treatment. Possibly, smart and ambitious students perform better at school, and high performance and ambition prompt parents to migrate to support children's schooling. Another scenario is that poverty may hinder children's educational achievement while driving parents to migrate. The endogeneity leads to selection bias. To tackle this bias, an applicable technique is matching. It aims to make two groups comparable by pairing a unit in the treatment group with one or several units in the control group based on observable covariates (Costalli \& Negri, 2021).

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-botev2019}{}}%
Botev, J., Égert, B., Smidova, Z., \& Turner, D. (2019). \emph{A new macroeconomic measure of human capital with strong empirical links to productivity}. OECD Economics Department Working Papers. \url{https://doi.org/10.1787/d12d7305-en}

>>>>

The internal migration stemmed from the neoliberal economic regime. Scholars like Duckett (2020) and Zhang (2018) argued that China conforms to the neoliberal trend, extending the logic of the market to other domains of human life. In the 1980s, the ruling communist party ditched the planned economy, initiated market reforms, and opened up the country. Since then, China has integrated in the global economy and recorded substantial growth. The economic reform and globalization have enriched the coastal cities like Shanghai and Shenzhen. The rural hinterlands, however, have lagged behind. Simultaneously, many people have left the countryside for the city. The regional inequality needs social policy to amend. But China's welfare program, as Qian (2021) argued, is fragmented and stratified. People who reap the greatest benefits are public servants, urban dwellers, and formally employed workers. As rural migrants arrive in the city, they find themselves with limited access to social welfare, such as health care and children's education (Qian, 2021). The welfare system's exclusion of rural migrants means that the state shifts its responsibility to individuals, resembling the neoliberal governmentality (Zhan, 2020).

@@

In other words, the hukou system binds one's citizenship rights with one's place of registration. In theory, Chinese citizens can migrate freely within the country. In practice, they cannot change hukou freely. When rural residents move to the city, it takes months -- if not years -- to apply for a new hukou. Before receiving urban hukou, they can hardly send their children into public schools in the host city (Wan \& Vickers, 2022; Zhang, 2018). Migrant workers have organized petitions, strikes, and protests. These actions often fell on deaf ears and sometimes even ended with police crackdown (Chan, 2019; Göbel, 2021). Facing financial and policy constraints, many migrant laborers have to leave children in their hometowns. Only during the winter and summer breaks can these children reunite with parents.

@@

Social scientists have researched extensively about the education of Chinese left-behind children, but the current literature still has three major shortcomings. The first shortcoming lies in sampling. For many studies, the geographic coverage is limited to one to a few counties; The participants are children left behind in rural areas but not those in urban areas (see Hu, 2019; Zhou et al., 2014). Alpermann (2020) pointed that the Chinese government reinforces ``an urban hierarchy'' besides the rural-urban divide. In the hierarchy's bottom are small- and medium-sized cities in the inland. They also see parents migrating to metropolises and leaving children behind (S. Wang, 2023). The National Bureau of Statistics noted that the number of urban left-behind children have been growing, and the trend is expected to continue (NBS et al., 2017). The second shortcoming is the usage of cross-sectional data. Wassink and Viera (2021) argued that when cross-sectional surveys merely focus on single time points, they fail to capture the dynamic nature of migratory journey. Research results would be under- or over-estimated ``depending on the timing of data collection'' (p.~1766). The third shortcoming lies in the measurement of outcome. Chang et al. (2019) and L. Wang et al. (2019) measured students' educational performance with the standardized test score of only one academic subject, namely mathematics. To tackle these limitations, I adopted a nationally representative two-wave panel dataset. My sample includes urban left-behind children as well. I included the test results of multiple subjects to measure comprehensive academic achievement.

@@

Between parental migration and children's education, the mechanisms trace to two theoretical frameworks. One is the resource generation model. It holds that parental migration brings forth greater resources in the form of economic capital (Brauw, 2019). Economic capital is, in Bourdieu's words, ``immediately and directly convertible into money and may be institutionalized in the form of property rights'' (2002, p. 16). Economic resources can help individuals acquire human capital, which is the stock of ``knowledge, skills, and health that people accumulate over their lives'' (World Bank, 2020, p. 1). Yang (2008) noted that in the Philippines, a sharp depreciation of peso increased the amount of remittances from migrants working abroad. This subsequently boosted children's school enrollment and educational spending. Vikram (2021) performed propensity score matching and fixed-effect regression on a two-wave panel dataset featuring Indian children aged 8-11. She showed that fathers' migration was linked with increases in left-behind children's reading capability and educational expenditure (p.~991). In China, Gu (2022) interviewed migrant workers and their family members. She found that they aspire to move up the social ladder; Such aspiration translates into ``intensive and all-out investment in children's education'' (p.~526). Oftentimes, labor migration presented the only way to sustain their children's education and livelihood. Peng (2018) interviewed first-generation migrant mothers whose initial migration dated back to the 1990s. To them, ``good mothering'' primarily meant offering economic support for their children in education, which can help their children climb the ladder of success.

The other theoretical model is the family disruption model. Children flourish under parental nurturing care. This term refers to the ``stable environments that promote health and adequate nutrition, protect from threats \ldots{} opportunities for learning,'' and ``responsive, emotionally supportive and developmentally enriching relationships'' (Black et al., 2021, p. 1). Parental migration decreases nurturing care, as characterized by less homework supervision and emotional support by parents. Hong talked to a group of left-behind children at a rural school (Hong \& Fuller, 2019). These students expressed the utter sense of loneliness ``with no advocate, no support and with no one looking out for them'' (p.~13). Many of them felt indifferent to grades and disengaged with school. Some even longed to quit school to pursue a free, independent life (p.~14). Murphy (2022) talked to left-behind adolescents at a Chinese village. Some of them ``embraced the `left-behind' label'' because it captured their need for more parental care (p.~185), ``especially companionship and affection in daily life'' (p.~186). According to C. Chen (2023), caregiver-child interactions are infrequent and unstable for left-behind children. She showed that the loss in such interactions ``mediates almost all of the adverse effects of parental absence.'' Her conclusion was based on structural equation modeling using a cross-sectional survey with a sample of Chinese children aged three to fifteen.

@@

I obtained the data from the China Education Panel Survey (CEPS) (NSRC, n.d.). It is a nationally representative, school-based survey featuring junior high school students. Currently, two waves of data are available. The CEPS team operates under the National Survey Research Center at the Renmin University of China. The survey team adopted multi-stage stratified probability proportional to size sampling and administered a paper-and-pencil questionnaire to each participant. The wave one baseline survey took place in the 2013-2014 academic year with a sample size of about twenty thousand students in the seventh and ninth grades. They were nested in 438 classrooms of 112 schools in 28 urban districts or rural counties (hereafter `counties'). The wave two follow-up survey in the 2014-2015 academic year lost track of the ninth grade cohort and only surveyed 10279 of the seventh grade cohort. I further restricted the sample to students living with both parents at the original domicile (place of hukou registry) at the baseline survey. To tackle missing values, I adopted list-wise deletion for the dependent, independent, and control variables. The final sample size amounts to 4114.

@@

The lists of covariates are in Table~\ref{tab:tab-ctvar-dit} -- Table~\ref{tab:tab-ctvar-msr} in the Results section. At the children's level, I controlled for age, sex, self-rated health, ethnicity, hukou type, number of siblings, whether the child attends boarding school, and whether the child had skipped grade or repeated grade. Age, sex, and health are associated with educational performance (Gu \& Yeung, 2021). Regarding ethnicity and hukou status, Han Chinese and urban hukou holders outperform ethnic minorities and rural hukou holders at school (Hasmath, 2022; Huang et al., 2020; D. Xu et al., 2018). I controlled for boarding at school because research results vary regarding the effect of boarding at school on academic performance (S. Guo et al., 2021; M. Liu \& Villa, 2020). I also included the number of siblings as a covariate, considering the quantity-quality trade-off in multi-child family (R. Guo et al., 2022). At the household level, I controlled for the levels of education of father and mother, whether the parent has a white collar job, whether the household has internet access, the perceived household financial conditions, and the number of extracurricular books at home. These variables indicate family socioeconomic status, which is related to children's educational achievement (J. Liu et al., 2020; Young \& Hannum, 2018). Finally, I controlled for the level of education that the child aspires to achieve and the parental expectation on the child's level of education. Children's academic achievement, their educational aspiration, and their parents' expectations are interrelated (Chen et al., 2023; Xu \& Montgomery, 2021). All control variables were measured in the baseline survey.

@@

I built a two-way fixed-effect regression model in a difference-in-differences (DID) design, utilizing parental migration as the treatment (Bai et al., 2018). The treatment group consists of students left behind by migrating parents at a certain point of time between the two waves of survey, and the control group includes those who do not. I compared the treatment group's pre- and post-treatment outcomes in relation to the control group.

Below is my first version of DID model: \[\Delta score_{i,j} = \alpha + \beta \cdot migr_{i} + \gamma \cdot score_{i;1} + \lambda \cdot S_{j} + \varepsilon_{i,j}\] It is unrestricted because it does not restrict on the coefficient associated with the baseline scores. The model is unadjusted as it does not adjust for additional covariates. Theoretically, it is unnecessary to include covariates that vary over group but remain constant over time; They would cancel out in the two-way fixed-effect model (Huntington-Klein, 2022). For student \(i\) in school \(j\), \(\Delta score_{i,j}\) is the change in test score between the two waves of surveys, \(migr_{i}\) is the treatment dummy variable, \(score_{i,j;1}\) is the test score at baseline, and \(S_{j}\) is the school fixed-effect. School fixed-effect can account for differences due to time-invariant factors at the school and county levels, as students nested in one school are also in the same county. The standard errors are clustered at the school level. An alternative is to use a class fixed-effect to control for time-invariant factors at the classroom level and above. However, about one-sixth of the students were reassigned into new classes at the end of baseline academic year. As the reassignment of classroom will complicate the model, I ruled out the classroom fixed-effect.

@@

I adopted \texttt{R} (version 4.2.2) for all the analyses (R Core Team, 2022). This involved the following packages: \texttt{tidyverse} (version 2.0.0) and \texttt{haven} (version 2.5.0) for data cleaning (Wickham et al., 2019, 2023); \texttt{fixest} (version 0.11.1) for econometric modeling (Berge et al., 2023); \texttt{MatchIt} (version 4.5.1) and \texttt{cobalt} (version 4.3.2) for matching (Greifer, 2022; Ho et al., 2023, 2011); \texttt{modelsummary} (version 1.3.0), \texttt{vtable} (version 1.4.1), and \texttt{kableExtra} (version 1.3.4) for presenting the results (Arel-Bundock, 2022; Huntington-Klein, 2023; Zhu, 2021); and \texttt{papaja} (version 0.1.1) for typesetting (Aust \& Barth, 2022).

@@

Table~\ref{tab:tab-idpvar} illustrates the independent variables, namely the migration status reported in the wave two survey. All 4114 children were living with parents in one household in baseline, but about sixteen percent saw at least one parent migrated between the two waves of survey. I call these students ``newly left-behind children'' and their households ``new migration households.'' In these new migration households, only-mother migration and only-father migration each accounted for two out of five cases. The remaining one-fifth were both-parents migration.

@@

\begin{longtable}[]{@{}lll@{}}
\caption{\label{tab:tab-idpvar}Independent variables by parental migration}\tabularnewline
\toprule()
Variable & N & Percent \\
\midrule()
\endfirsthead
\toprule()
Variable & N & Percent \\
\midrule()
\endhead
Parental Migration & & \\
\ldots{} No & 3465 & 84\% \\
\ldots{} Yes & 649 & 16\% \\
Parental Migration Type & & \\
\ldots{} Both Parents at Home & 3465 & 84\% \\
\ldots{} Both Parents Migrated & 132 & 3\% \\
\ldots{} Only Father Migrated & 256 & 6\% \\
\ldots{} Only Mother Migrated & 261 & 6\% \\
\bottomrule()
\end{longtable}

@@

\caption{\label{tab:tab-dpvar}Dependent variables by parental migration}
\centering
\begin{tabular}[t]{llllll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Parental Migration} \\
\cmidrule(l{3pt}r{3pt}){2-5}
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{No = 3465} & \multicolumn{2}{c}{Yes = 649} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
Variable & Mean & Sd & Mean & Sd & Test\\
\midrule
Cognitive test score difference & 11 & 20 & 9.6 & 21 & p=0.027*\\
\hspace{1em}Wave 1 score & 57 & 17 & 56 & 17 & p=0.312\\
\hspace{1em}Wave 2 score & 68 & 18 & 66 & 20 & p=0.001**\\
Academic exam score difference & -2.8 & 11 & -4.1 & 12 & p=0.007**\\
\hspace{1em}Wave 1 score & 71 & 16 & 70 & 17 & p=0.24\\
\addlinespace
\hspace{1em}Wave 2 score & 68 & 18 & 66 & 20 & p=0.008**\\
\bottomrule
\multicolumn{6}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{6}{l}{\rule{0pt}{1em}* p < 0.05, ** p < 0.01, *** p < 0.001}\\
\end{tabular}
\end{table}

@@

\caption{\label{tab:tab-ctvar-dit}Dichotomous control variables by parental migration}
\centering
\begin{tabular}[t]{llllll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Parental Migration} \\
\cmidrule(l{3pt}r{3pt}){2-5}
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{No = 3465} & \multicolumn{2}{c}{Yes = 649} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
Variable & N & Percent & N & Percent & Test\\
\midrule
child.is.girl &  &  &  &  & X2=12.716***\\
... 0 & 1660 & 48\% & 361 & 56\% & \\
... 1 & 1805 & 52\% & 288 & 44\% & \\
child.has.rural.hukou &  &  &  &  & X2=1.106\\
... 0 & 1912 & 55\% & 343 & 53\% & \\
\addlinespace
... 1 & 1553 & 45\% & 306 & 47\% & \\
child.in.boarding.school &  &  &  &  & X2=3.876*\\
... 0 & 2500 & 72\% & 443 & 68\% & \\
... 1 & 965 & 28\% & 206 & 32\% & \\
child.is.ethnic.minority &  &  &  &  & X2=0.585\\
\addlinespace
... 0 & 3230 & 93\% & 599 & 92\% & \\
... 1 & 235 & 7\% & 50 & 8\% & \\
child.had.skipped.grade &  &  &  &  & X2=0.054\\
... 0 & 3429 & 99\% & 641 & 99\% & \\
... 1 & 36 & 1\% & 8 & 1\% & \\
\addlinespace
child.had.repeated.grade &  &  &  &  & X2=12.906***\\
... 0 & 3160 & 91\% & 562 & 87\% & \\
... 1 & 305 & 9\% & 87 & 13\% & \\
child.went.to.preschool &  &  &  &  & X2=0.826\\
... 0 & 530 & 15\% & 109 & 17\% & \\
\addlinespace
... 1 & 2935 & 85\% & 540 & 83\% & \\
parent.has.white.collar.job &  &  &  &  & X2=0.001\\
... 0 & 2739 & 79\% & 512 & 79\% & \\
... 1 & 726 & 21\% & 137 & 21\% & \\
home.has.internet &  &  &  &  & X2=13.27***\\
\addlinespace
... 0 & 1064 & 31\% & 247 & 38\% & \\
... 1 & 2401 & 69\% & 402 & 62\% & \\
\bottomrule
\multicolumn{6}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{6}{l}{\rule{0pt}{1em}0 = No, 1 = Yes; * p < 0.05, ** p < 0.01, *** p < 0.001}\\
\end{tabular}
\end{table}

@@

\caption{\label{tab:tab-ctvar-cnt}Categorical and continuous control variables by parental migration}
\centering
\begin{tabular}[t]{llllll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Parental Migration} \\
\cmidrule(l{3pt}r{3pt}){2-5}
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{No = 3465} & \multicolumn{2}{c}{Yes = 649} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
Variable & Mean & Sd & Mean & Sd & Test\\
\midrule
child.age & 12 & 0.63 & 13 & 0.73 & F=5.008*\\
child.number.of.sibling & 0.58 & 0.74 & 0.66 & 0.82 & F=5.242*\\
child.health & 4.2 & 0.87 & 4.1 & 0.89 & F=0.575\\
child.education.aspiration & 4 & 1.1 & 3.9 & 1.2 & F=0.496\\
mother.education.level & 1.5 & 1.2 & 1.6 & 1.3 & F=0.48\\
\addlinespace
father.education.level & 1.7 & 1.2 & 1.8 & 1.2 & F=0.413\\
parent.expectation & 4.2 & 0.91 & 4.1 & 0.95 & F=1.598\\
home.extracurricular.book & 3.4 & 1.2 & 3.3 & 1.3 & F=3.82\\
home.economic.resource & 2.9 & 0.55 & 2.8 & 0.57 & F=1.399\\
\bottomrule
\multicolumn{6}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{6}{l}{\rule{0pt}{1em}* p < 0.05, ** p < 0.01, *** p < 0.001}\\
\end{tabular}
\end{table}

@@

I first explain the results on children's cognitive test score. The unadjusted DID models show that newly left-behind adolescents are more likely to score lower in cognitive tests than do their non-left-behind counterparts (see Table 6 below). This lends support to Hypothesis 1a. Parental migration is linked to a 2.54 points decrease in cognitive test score. Hypothesis 2a is only partially supported. Compared with the non-left-behind peers, children with only father absent are associated with greater decrease (2.81 points) in cognitive test score than those with only mother absent (2.16 points). The adjusted DID models produce results similar to the unadjusted models, with one exception (see Table A1 in the appendix). That is, for the cognitive score, the both-parent-migrated coefficient's magnitude is lower than that of father-migrated.

@@

Parental migration is not a randomly assigned treatment. Possibly, smart and ambitious students perform better at school, and high performance and ambition prompt parents to migrate to support children's schooling. Another scenario is that poverty may hinder children's educational achievement while driving parents to migrate. The endogeneity leads to selection bias. To tackle this bias, an applicable technique is matching. It aims to make two groups comparable by pairing a unit in the treatment group with one or several units in the control group based on observable covariates (Sävje et al., 2021).

@@

\leavevmode\vadjust pre{\hypertarget{ref-alpermann2020}{}}%
Alpermann, B. (2020). {China}{'}s Rural{\textendash}Urban Transformation: New Forms of Inclusion and Exclusion. \emph{Journal of Current Chinese Affairs}, \emph{49}(3), 259--268. \url{https://doi.org/10.1177/18681026211004955}

@@
